---
title: "HW4 Rmarkdown and Litterfall Module Synthesis"
author: "Alexia Ford"
date: "2/26/2018"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
The NEON (National Ecological Observatory Network) project provides scientific real time data in hopes of understanding environmental change, informing natural resource decisions, and engaging the next generations of scientists with this data. The project is responsible for measuring the causes and effects of environmental change across the United States within 20 eco-climate domains that all represent different landforms, vegetation, climate, and ecosystem dynamics. The project is constantly collecting data from it's observatory's regional domain to provide data and statistically represent ecological, physical, and biological variability across the US (NEON data website). 

NEON has four different sites that the project collects data from that will be introduced in this project, these sites include: core terrestrial sites, core aquatic sites, relocatable terrestrial sites, and lastly relocatable aquatic sites (NEON data website).In this analysis we will specifically be looking at: the Barlett Experimental Forest (relocatable terrestrial), Guanica Forest Common Wealth of Puerto Rico (core terrestrial), Harvard University Forest (core terrestrial), Santa Rita Experimental Range University of Arizona (core terrestrial), Steigerwaldt Land Services (relocatable terrestrial), Treehaven University of Wisconsin (relocatable terrestrial), and University of Notre Dame Environmental Research Center (core terrestrial).

The site I will specifically be focused on for this project is the "STEI" site, other wise known as Steigerwaldt Land Services. This site is a relocatable terrestrial site located near the Great Lakes in Tomahawk, Wisconsin. It is approxamitely located 1 mile north of NEONs other Treehaven site. Steigerwaldt has a specific vegetation consisting of the most abundant tree species with young, even aged stand such as: aspen (Populus tremuloides), red maple (Acer rubrum), and balsam fir (Abies balsamea). The stand was cut in 2005 and was scattered with oak and hemlock and roads were planted with clover for wildlife food for deer and ruffed grouse. 

In addition to vegetation the area has a disinguishable climate. "STEI" has endured salvage timber projects following ice and hail damage, windstorms, tornadoes, and spruce decline. Steigerwaldt is typically faces a humid continental climate and experiences no dry seasons and warm summers from mid-May to mid-September with average temperatures in July around 79 F. Contrastingly the site faces winters lasting from Novemeber to March with temperature arounds 4 F in January. The sites annual average temperature is 41.6 F.  

# Methods

Describe briefly where the data came from, including how they were collected in the field (information is in the pdf file in the data folder), and how you are analyzing them (e.g. using R to visualize the results and an ANOVA to test for significant differences).

# Results
Read in Litterfall data set:
```{r load-in-data}
# read in the literfall data from a csv
literfall_data <- read.csv("data/ltr_massdata.csv",
                           header = TRUE)
```

Data Set showing the dry mass of Branches and Twigs at NEON sites:
```{r subset-out-twigs/branches-data}
# just look at twigs and branches dry mass
branches_data <- subset(literfall_data, 
                        functionalGroup == "Twigs/branches")
```

```{r subset-out-STEI-twigs/branches-data}
#just look at the "STEI" neon site
STEI_branches_data <- subset(branches_data,
                    siteID == "STEI")
```

Scatter plot of the Branches/Twigs mass at STEI NEON site:
```{r scatterplot-for-STEI-site}
# make a scatterplot of branches and twigs dry mass at the STEI NEON site
pdf("output/STEI_branches_mass_over_time_scatterplot.pdf")
plot(x = as.POSIXct(STEI_branches_data$collectDate),
    y = STEI_branches_data$dryMass,
    main = "Dry mass of branches fallen in Steigerwaldt Land Service",
    xlab = "Date",
    ylab = "Dry mass of banches in grams",
    pch = 16)
dev.off()
```

```{r install-lintr}
install.packages("devtools")
devtools::install_github("jimhester/lintr")
```

```{r check code}
lintr::lint(filename = "Analysis.Rmd")
```

Add a number of code chunks in the Results section. These should read in, subset and plot the data as needed (no need to save it to pdf, since it will be put into the rendered docx when you click 'knit'), and, for any boxplots you make, run an appropriate statistical test on your dataset.

# Discussion

Discuss the results you found. In the context of what you know about the site/sites and the predominant types of vegatation in those locations. What do the results indicate about the flow of C and N from the autotrophic biota and into the soil in comparison to what occurs at other sites and over time?
