---
title: "HW4 Rmarkdown and Litterfall Module Synthesis"
author: "Alexia Ford"
date: "2/26/2018"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Delete this text, and add 3-4 paragraphs of introduction. This should include background on the NEON project, the sets of sites you are comparing, and the focal site you select for focused temporal analysis.

# Methods

Describe briefly where the data came from, including how they were collected in the field (information is in the pdf file in the data folder), and how you are analyzing them (e.g. using R to visualize the results and an ANOVA to test for significant differences).

# Results
Read in Litterfall data set:
```{r load-in-data}
# read in the literfall data from a csv
literfall_data <- read.csv("data/ltr_massdata.csv",
                           header = TRUE)
```

Data Set showing the dry mass of Branches and Twigs at NEON sites:
```{r subset-out-twigs/branches-data}
# just look at twigs and branches
branches_data <- subset(literfall_data, 
                        functionalGroup == "Twigs/branches")
```

```{r subset-out-STEI-twigs/branches-data}
#just look at the "STEI" neon site
STEI_branches_data <- subset(branches_data,
                    siteID == "STEI")
```

Scatter plot of the Branches/Twigs mass at STEI NEON site:
```{r scatterplot-for-STEI-site}
# make a scatterplot of branches and twigs litter at the STEI NEON site
pdf("output/STEI_branches_mass_over_time_scatterplot.pdf")
plot(x = as.POSIXct(STEI_branches_data$collectDate),
    y = STEI_branches_data$dryMass,
    main = "Dry mass of branches fallen in Steigerwaldt Land Service",
    xlab = "Date",
    ylab = "Dry mass of banches in grams",
    pch = 16)
dev.off()
```

```{r install-lintr}
install.packages("devtools")
devtools::install_github("jimhester/lintr")
```

```{r check code}
lintr::lint(filename = "Analysis.Rmd")
```

Add a number of code chunks in the Results section. These should read in, subset and plot the data as needed (no need to save it to pdf, since it will be put into the rendered docx when you click 'knit'), and, for any boxplots you make, run an appropriate statistical test on your dataset.

# Discussion

Discuss the results you found. In the context of what you know about the site/sites and the predominant types of vegatation in those locations. What do the results indicate about the flow of C and N from the autotrophic biota and into the soil in comparison to what occurs at other sites and over time?
